---
title: Multinomial Regression using Hyperparameter Optimization with SVM
author: "Niharika Sharma"
date: "September 19, 2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---



```{r setup, include=FALSE}
Sys.setenv(PATH = paste(Sys.getenv("PATH"), "C:\\Users\\NIHA\\AppData\\Local\\Programs\\MikTex 2.9\\miktex\\bin", sep=.Platform$path.sep))
knitr::opts_chunk$set(echo = FALSE)

#install.packages('tinytex')
#tinytex::install_tinytex()
```



```{r loadpackages, warning=FALSE}
pacman::p_load(e1071, tidyverse, caret, rmarkdown, corrplot, readxl, ModelMetrics)
theme_set(theme_classic())
```

## Defining data
Certain columns have been deleted due to redundancy issues.
The List of the deleted columns:
PriceCH
PriceMM
Store7
ListPriceDiff
STORE

More crucial attribute than the prices of both these brands is the price difference between both these brands.
We already have the Store ID in the dataset so STORE and Store7 need not be included in the dataset.
List Price difference is a redundant attribute as we already have the Sales Price Difference in the data.

Removing such redundant variables in the dataset and keeping more relevant attributes will help us avoid the overfitting of the model.

Creating a Partition for the train and test datasets.

```{r dataload , echo=TRUE}
juice <- read_csv("juice.csv")

drops <- c("PriceCH", "PriceMM" ,"Store7", "ListPriceDiff",	"STORE")
juice <- select(juice,-drops)

juice$Purchase <- as.factor(juice$Purchase)

set.seed(123)
train.index <- createDataPartition(juice$Purchase , p=0.8, list=FALSE)
juice_train <- juice[train.index,]
juice_test <- juice[-train.index,]
```


## Linear SVM Model
```{r LinearSVM, echo=TRUE}
set.seed(123)
svm_lnr <- svm(Purchase~., data=juice_train, kernel="linear", cost=0.01)
summary(svm_lnr)

pred_lnr1 <- predict( svm_lnr, juice_train)
pred_lnr2 <- predict( svm_lnr, juice_test)

confmatrix_lnr1 <- table(Actual = juice_train$Purchase, Predicted = pred_lnr1)
confmatrix_lnr1

confmatrix_lnr2 <- table(Actual = juice_test$Purchase, Predicted = pred_lnr2)
confmatrix_lnr2

```
Train Accuracy Score - `r sum(diag(confmatrix_lnr1))/sum(confmatrix_lnr1)  ` 

Test Accuracy Score - `r sum(diag(confmatrix_lnr2))/sum(confmatrix_lnr2)  `

Train Error Rate - `r 1-sum(diag(confmatrix_lnr1))/sum(confmatrix_lnr1)  ` 

Test Error Rate - `r 1-sum(diag(confmatrix_lnr2))/sum(confmatrix_lnr2)  `


## Grid Search for Linear SVM using tune function

```{r Lnr_SVM_Tune  , echo=TRUE}
cost_arr=seq(0.01, 10, by=0.1)
set.seed(123)
tune_lnr <- tune(svm, Purchase~., data = juice_train, kernel="linear",
                 ranges = list(cost=cost_arr))

summary(tune_lnr)

plot(tune_lnr)


pred_lnr1 <- predict( tune_lnr$best.model, juice_train)
pred_lnr2 <- predict( tune_lnr$best.model, juice_test)

confmatrix_lnr1 <- table(Actual = juice_train$Purchase, Predicted = pred_lnr1)
confmatrix_lnr1

confmatrix_lnr2 <- table(Actual = juice_test$Purchase, Predicted = pred_lnr2)
confmatrix_lnr2


```

Train Accuracy Score - `r sum(diag(confmatrix_lnr1))/sum(confmatrix_lnr1)  `

Test Accuracy Score - `r sum(diag(confmatrix_lnr2))/sum(confmatrix_lnr2)  ` 

Train Error Rate - `r 1-sum(diag(confmatrix_lnr1))/sum(confmatrix_lnr1)  ` 

Test Error Rate - `r 1-sum(diag(confmatrix_lnr2))/sum(confmatrix_lnr2)  `

##############################################################################

## RBF Model 

```{r RBF  , echo=TRUE}
set.seed(123)
svm_rbf <- svm(Purchase~., data=juice_train, cost=0.01)

summary(svm_rbf)

pred_rbf1 <- predict( svm_rbf, juice_train)
pred_rbf2 <- predict( svm_rbf, juice_test)

confmatrix_rbf1 <- table(Actual = juice_train$Purchase, Predicted = pred_rbf1)
confmatrix_rbf1

confmatrix_rbf2 <- table(Actual = juice_test$Purchase, Predicted = pred_rbf2)
confmatrix_rbf2
```
Train Accuracy Score - `r sum(diag(confmatrix_rbf1))/sum(confmatrix_rbf1)  `

Test Accuracy Score - `r sum(diag(confmatrix_rbf2))/sum(confmatrix_rbf2)  ` 

Train Error Rate - `r 1-sum(diag(confmatrix_rbf1))/sum(confmatrix_rbf1)  ` 

Test Error Rate - `r 1-sum(diag(confmatrix_rbf2))/sum(confmatrix_rbf2)  `

## Tuned RBF Model

```{r RBF_SVM_Tune  , echo=TRUE}
set.seed(123)
tune_rbf <- tune(svm, Purchase~., data = juice_train,
                 ranges = list(cost=cost_arr))

summary(tune_rbf)

plot(tune_rbf)

pred_rbf1 <- predict( tune_rbf$best.model, juice_train)
pred_rbf2 <- predict( tune_rbf$best.model, juice_test)

confmatrix_rbf1 <- table(Actual = juice_train$Purchase, Predicted = pred_rbf1)
confmatrix_rbf1

confmatrix_rbf2 <- table(Actual = juice_test$Purchase, Predicted = pred_rbf2)
confmatrix_rbf2
```
Train Accuracy Score - `r sum(diag(confmatrix_rbf1))/sum(confmatrix_rbf1)  `

Test Accuracy Score - `r sum(diag(confmatrix_rbf2))/sum(confmatrix_rbf2)  ` 

Train Error Rate - `r 1-sum(diag(confmatrix_rbf1))/sum(confmatrix_rbf1)  ` 

Test Error Rate - `r 1-sum(diag(confmatrix_rbf2))/sum(confmatrix_rbf2)  `

####################################################################

## Polynomial Model

```{r Poly  , echo=TRUE}
set.seed(123)
svm_poly <- svm(Purchase~., data=juice_train, kernel ="poly", cost=0.01, degree=2)

summary(svm_poly)

pred_poly1 <- predict(svm_poly, juice_train)
pred_poly2 <- predict(svm_poly, juice_test)

confmatrix_poly1 <- table(Actual = juice_train$Purchase, Predicted = pred_poly1)
confmatrix_poly1

confmatrix_poly2 <- table(Actual = juice_test$Purchase, Predicted = pred_poly2)
confmatrix_poly2
```
Train Accuracy Score - `r sum(diag(confmatrix_poly1))/sum(confmatrix_poly1)  `

Test Accuracy Score - `r sum(diag(confmatrix_poly2))/sum(confmatrix_poly2)  ` 

Train Error Rate - `r 1-sum(diag(confmatrix_poly1))/sum(confmatrix_poly1)  ` 

Test Error Rate - `r 1-sum(diag(confmatrix_poly2))/sum(confmatrix_poly2)  `

```{r Poly_SVM_Tune  , echo=TRUE}
set.seed(123)
tune_poly <- tune(svm, Purchase~., data = juice_train, kernel="poly", degree=2,
                 ranges = list(cost=cost_arr))

summary(tune_poly)
plot(tune_poly)

pred_poly1 <- predict( tune_poly$best.model, juice_train)
pred_poly2 <- predict( tune_poly$best.model, juice_test)

confmatrix_poly1 <- table(Actual = juice_train$Purchase, Predicted = pred_poly1)
confmatrix_poly1

confmatrix_poly2 <- table(Actual = juice_test$Purchase, Predicted = pred_poly2)
confmatrix_poly2


```

Train Accuracy Score - `r sum(diag(confmatrix_poly1))/sum(confmatrix_poly1)  ` 

Test Accuracy Score - `r sum(diag(confmatrix_poly2))/sum(confmatrix_poly2)  ` 

Train Error Rate - `r 1-sum(diag(confmatrix_poly1))/sum(confmatrix_poly1)  ` 

Test Error Rate - `r 1-sum(diag(confmatrix_poly2))/sum(confmatrix_poly2)  `

## Conclusion


After comparing the Train and Test Scores for all the models:


<u>Basic Models</u>


SVM with Linear Kernel is the best in case of Basic Models with the least Error scores for both Train and Test datasets.




<u>Tuned Models</u>


For both RBF and Linear Kernels, the cost parameter for the best model is 0.31 and the scores are almost equal.
For Polynomial, the cost parameter is 9.61 but the model isn't as good.  

Taking the Accuracy rate and Error Rate into Consideration, both tuned models - SVM with Kernel RBF and Kernel Linear are good but RBF is slighlty better as there are lower chances of Overfitting.


#####################################################################################